---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# percentify

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/percentify)](https://cran.r-project.org/package=percentify)
[![Travis build status](https://travis-ci.org/EmilHvitfeldt/percentify.svg?branch=master)](https://travis-ci.org/EmilHvitfeldt/percentify)
[![Codecov test coverage](https://codecov.io/gh/EmilHvitfeldt/percentify/branch/master/graph/badge.svg)](https://codecov.io/gh/EmilHvitfeldt/percentify?branch=master)
<!-- badges: end -->

The goal of percentify is to create virtual groups on top of a `tibble` or `grouped_df` to allow calculation within percentile ranges of a variable on the whole dataset.  You can then efficiently perform various dplyr operations on this resampled_df, like: `summarise()`, `do()` and `group_map()`.

## Installation

You can install the developmental version of percentify from Github with:

``` r
devtools::install_github("EmilHvitfeldt/percentify")
```

## Example

Imagine we want to do some summary statistics at the different percentile ranges of price in diamonds. We start by using `percentify_cut` to created a `percentiled_df` on price with splits at `20%`, `60%`, `80%`, `90%` and `95%`.

```{r message=FALSE}
library(percentify)
library(tidyverse)
```

```{r}
diamonds_price <- percentify_cut(diamonds, price, c(0.2, 0.6, 0.8, 0.9, 0.95))

diamonds_price
```

We can then use this grouped data.frame with `summarise` to canculate statistics within each range.

```{r}
summarise(diamonds_price,
          mean_carat = mean(carat),
          procent_ideal = mean(cut == "Ideal"),
          mean_x = mean(x),
          n_obs = n())
```

Using `collect` from dplyr will materialize the groups so they can be used for plotting or other calculations.

```{r}
diamonds_price %>%
  collect() %>%
  ggplot(aes(x, fill = .percentile_price)) +
  geom_histogram(bins = 100)
```

## PLotting function

The resulting grouped data.frame have `ggplot2::autoplot()` methods to vizualize the the percentile ranges.

```{r}
percentify_random(diamonds, price, 0.2, 25) %>%
  autoplot()
```


## Inspiration

The underlaying code for this package is inspired by the work done by [Davis Vaughan](https://twitter.com/dvaughan32) in [strapgod](https://github.com/DavisVaughan/strapgod).

## Code of Conduct

Please note that the 'quansum' project is released with a [Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md). By contributing to this project, you agree to abide by its terms.
